---
title: "Vignette MorphoGraphX2R"
author: "Marion Louveaux"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette MorphoGraphX2R}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Librairies needed

```{r lib}
library(colorRamps)
library(MorphoGraphX2R)
library(magrittr)
library(dplyr)
library(plotly)
library(glue)
library(RColorBrewer)
```

## Loading data

All .ply demonstration data are stored in /inst/extdata.
```{r allData}
dirData <- "/Users/mlouveaux/Documents/Postdoc/Projects/MGX-R bridge/MorphoGraphX2R/inst/extdata"
```

We select one .ply as example.
```{r myData}
filePly <- paste0(dirData, "/normalMesh/2013-02-12_LTi6B_dis_A_T0_cells_minW1_normalMesh.ply")
fileCellGraph <- paste0(dirData, "/cellGraph/2013-02-12_LTi6B_dis_A_T0_cells_minW1_cellGraph.ply")

```


Note from 2017-12-26: read_ply_function is clean. I need now to test it on numerous meshes to make it as general as possible.


## Mesh and cell graph (e.g. cells and edges properties)

```{r readPly}
# N.B. to check mesh: put rgl viz out of modified_read.ply
myMesh <- modified_read.ply(file = filePly, ShowSpecimen = FALSE, addNormals = TRUE,
                               MatCol= 1, header_max = 30,
                               my_colors = c("#800000", "#FF0000", "#808000", "#FFFF00",
                                             "#008000", "#00FF00", "#008080", "#00FFFF",
                                             "#000080", "#0000FF", "#800080", "#FF00FF"))
```

To display the mesh with RGL
```{r}
# rgl::shade3d(myMesh)
```

To change mesh color and display the new color:
```{r modifColor, eval=FALSE}
myMesh$material$color <- myMesh$allColors$Col_signal
#rgl::shade3d(myMesh)
```

Information relative to cells and cell walls (resp. vertices and edges of the cell graph) are stored into two dataframes and combined into a list:
```{r readCellGraph, eval=FALSE}
myCellGraph <- MorphoGraphX2R::modified_read.cellGraph(fileCellGraph = fileCellGraph, header_max = 30)
```



## Displaying the mesh with plotly

```{r}
# meshCellcenter <- myCellGraph$vertices[,c("label","x", "y", "z")]
meshCellcenter <- NULL
```

To display the mesh with cell labels:
```{r map3D, eval=FALSE}
# add control on the title of the plot
p1 <- plotlyMesh(meshExample = myMesh,
           meshColors = myMesh$allColors$Col_label,
           meshCellcenter = meshCellcenter)
p1
```

Without the cell labels:
```{r map3D2, eval=FALSE}
# add control on the level of opacity
plotlyMesh(meshExample = myMesh,
           meshColors = NULL,
           meshCellcenter = meshCellcenter)
```



## Displaying specific info from cellGraph on the mesh

### Scalar info (area, growth rate...)
```{r heatmap3D, eval=FALSE}
p2 <- plotlyMesh(meshExample = myMesh, 
           meshColors = left_join(myMesh$it_label, myCellGraph$vertices) %>%
             select(., GeometryArea),
           meshCellcenter =  meshCellcenter
)
p2
```


### Vector/tensor info (principal dir of curv...)

```{r heatmap3D2, eval=FALSE}

p3 <- plotlyMesh(meshExample = myMesh, 
                meshColors = NULL,
                meshCellcenter = meshCellcenter)

# Find coords
find_coords <- function(cellcenter, tipUnit, d, scale = 1) {
  d <- d*scale
  # Distance
  tip1 <- d / (sqrt(sum(tipUnit^2))) * tipUnit + cellcenter
  tip2 <- (-d / (sqrt(sum(tipUnit^2))) * tipUnit) + cellcenter
  tibble(x1 = tip1[1], y1 = tip1[2], z1 = tip1[3],
         x2 = tip2[1], y2 = tip2[2], z2 = tip2[3])
}

# Test avec !!
tensor_name <- "CellAxisPDG"
grand.name <- glue("{tensor_name}.{c(1:3, 7)}")
petit.name <- glue("{tensor_name}.{c(4:6, 8)}")

tmp <- myCellGraph$vertices %>%
  rename(
    tipUnit.1 = !!grand.name[1],
    tipUnit.2 = !!grand.name[2],
    tipUnit.3 = !!grand.name[3],
    d = !!grand.name[4],
  ) %>%
  mutate(grand = purrr::pmap(list(x, y, z, tipUnit.1, tipUnit.2, tipUnit.3, d),
                             ~find_coords(cellcenter = c(..1, ..2, ..3),
                                          tipUnit = c(..4, ..5, ..6),
                                          d = ..7,
                                          scale = 150))) %>%
  tidyr::unnest(grand, .sep = ".") %>%
  select(-starts_with("tipUnit"), -d) %>%
  rename(
    tipUnit.1 = !!petit.name[1],
    tipUnit.2 = !!petit.name[2],
    tipUnit.3 = !!petit.name[3],
    d = !!petit.name[4],
  ) %>%
  mutate(petit = purrr::pmap(list(x, y, z, tipUnit.1, tipUnit.2, tipUnit.3, d),
                             ~find_coords(cellcenter = c(..1, ..2, ..3),
                                          tipUnit = c(..4, ..5, ..6),
                                          d = ..7,
                                          scale = 150))) %>%
  tidyr::unnest(petit, .sep = ".") %>%
  select(-starts_with("tipUnit"), -d) %>%
  inner_join(myCellGraph$vertices)

for (i in 1:nrow(tmp)){
  xcoord <- as.vector(t(select(tmp, x, grand.x1, grand.x2)[i,]))
  ycoord <- as.vector(t(select(tmp, y, grand.y1, grand.y2)[i,]))
  zcoord <- as.vector(t(select(tmp, z, grand.z1, grand.z2)[i,]))
  p3 <- add_trace(p3, x = xcoord, y = ycoord, z = zcoord,
                 line = list(color = 'rgb(0, 0, 0)', width = 6), mode = "lines",
                 type = "scatter3d", hoverinfo = 'none') %>%
    layout(showlegend = FALSE)
  
  xcoord <- as.vector(t(select(tmp, x, petit.x1, petit.x2)[i,]))
  ycoord <- as.vector(t(select(tmp, y, petit.y1, petit.y2)[i,]))
  zcoord <- as.vector(t(select(tmp, z, petit.z1, petit.z2)[i,]))
  p3 <- add_trace(p3, x = xcoord, y = ycoord, z = zcoord,
                 line = list(color = 'rgb(0, 0, 0)', width = 6), mode = "lines",
                 type = "scatter3d", hoverinfo = 'none') %>%
    layout(showlegend = FALSE)

}
p3
```


## Drawing cell contours

```{r}
vertexCellcenter <- purrr::map(1:ncol(myMesh$allColors$Col_label), ~ 
  myMesh$it[ which(myMesh$allColors$Col_label[,.x] == names(which(table(myMesh$allColors$Col_label[,.x]) == 1))), .x ]
)

edgesCoords <- purrr::map(1:ncol(myMesh$allColors$Col_label), ~ 
  myMesh$vb[,myMesh$it[ which(myMesh$allColors$Col_label[,.x] != names(which(table(myMesh$allColors$Col_label[,.x])==1))), .x]]
)

p <- plotlyMesh(meshExample = myMesh,
           meshColors = NULL)

for (i in 1:length(unique(edgesCoords))){
  edge_tmp <- unique(edgesCoords)[[i]]
  xcoord <- edge_tmp[1,]
  ycoord <- edge_tmp[2,]
  zcoord <- edge_tmp[3,]
  p <- add_trace(p, x = xcoord, y = ycoord, z = zcoord,
                 line = list(color = 'rgb(255, 255, 255)', width = 6), mode = "lines",
                 type = "scatter3d", hoverinfo = 'none') %>%
    layout(showlegend = FALSE)
}
p

```



